
\chapter{Conclusion}

In the thesis we presented the \ToolName tool that is able to compute
data lineage information for Java applications that use complex data processing
frameworks to access data.

A part of our work was to propose changes to the Symbolic analysis library
such that it is possible to easily add support for new frameworks as the library plugins.
Our main contribution is the implementation of such plugins
for selected frameworks - Spring JDBC, MyBatis and Kafka.
This demonstrates that such plugins can be used to provide
support for data lineage analysis of other Java data processing frameworks
with different approaches for accessing the data.

Each plugin should identify the sources and sinks of
data inside the corresponding framework and then the Symbolic analysis library takes care of
computing data flow in an analysed application.

We showed few limitations of our \ToolName tool. Many of them
are related to the static program analysis approach and the used
Symbolic analysis library.

However, the tests of our \ToolName tool show that the data flow is
computed in reasonable time and thereof can be used in practice.
The tests also create data lineage visualizations.
The visualizations contain the key information about
data sources and sinks and the flow between them.

In our work we focused on the basic features of selected data processing frameworks.
We avoid the advanced features, such as Kafka Streams or Kafka Connector.
We also did not try to provide the implementation for the the whole Spring JDBC Framework,
as it provides quite large possibilities for accessing the data in an SQL database.

In the future development of the \ToolName tool, we are planning to add support
for all the mentioned advanced features of the frameworks.
We will also develop plugins for new data processing frameworks
such as Hibernate and Apache Spark.


